Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-04 Master Thesis/synopsis.tex
Change phrasing of 2.5

\begin{lemma}
Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if $\dim k(v)=0$.
\end{lemma}

@@

\begin{lemma}\label{2.5}
Let $K$ be henselian with respect to $w\in|\operatorname{Spv}(K)|$. Set $k=k(w)$ and $\Gamma = \Gamma_w$. Then $k$ is an algebraic extension of a finite field $\mathbb{F}_p$. Let $\ell\neq p$ be a prime and assume $\mu_\ell\subset K$. Then the following conditions are equivalent:
\begin{enumerate}[(i)]
\item $\mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})=\mathbb{Z}/\ell\mathbb{Z}$
\item $\ell^\infty\nmid[k:\mathbb{F}_p]$ and $\dim_{\mathbb{F}_\ell}(\Gamma/\ell\Gamma) = 2$
\end{enumerate}
Otherwise we have $\mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z})=0$. If $L/K$ is an algebraic extension and (i) holds, we have:
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K]\\
0, & \text{otherwise}
\end{cases} \]
\end{lemma}

@@

Let $G_K(\ell)$ be the maximal pro-$\ell$-quotient of $G_K$. Since $\mu_\ell\subset K$, the action of $G_k$ on the pro-$\ell$-Sylow group of the inertia group of $w$ factors through $G_k(\ell)$. We see that the canonical projection $G_K\to G_K(\ell)$ is an isomorphism. Therefore we have:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z}) \cong\mathrm{H}^3(G_K(\ell),\mathbb{Z}/\ell) \cong\mathrm{H}^3(G_\ell,\mathbb{Z}/\ell\mathbb{Z}) \]
Now the equivalence of (i) and (ii) follows from [Se], Ch. I, §4.

>>>>

\begin{lemma}\label{2.2-prepre}
Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if it has no non-trivial specializations or iff $\dim k(v)=0$, i.e. $k(v)$ has positive characteristic and is algebraic over its prime field.
\end{lemma}

@@

\begin{lemma}\label{2.5}
Let $K$ be henselian with respect to $w\in|\operatorname{Spv}(K)|$. Set $k=k(w)$ and $\Gamma = \Gamma_w$. Then $k$ is an algebraic extension of a finite field $\mathbb{F}_p$ by Lemma \ref{2.2-prepre}. Let $\ell\neq p$ be a prime and assume $\mu_\ell\subset K$. Then:
\[ \mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [k:\mathbb{F}_p] \text{ and } \dim_{\mathbb{F}_\ell}(\Gamma / \ell\Gamma) = 2\\
0, & \text{otherwise}
\end{cases} \]
If $L/K$ is an algebraic extension and $\mathrm{H}^3(K, \mathbb{Z}/\ell\mathbb{Z}) = \mathbb{Z}/\ell\mathbb{Z}$ holds, we have:
\[ \mathrm{H}^3(L,\mathbb{Z}/\ell\mathbb{Z}) = \begin{cases}
\mathbb{Z}/\ell\mathbb{Z}, & \ell^\infty\nmid [L:K]\\
0, & \text{otherwise}
\end{cases} \]
\end{lemma}

@@

Let $G_K(\ell)$ be the maximal pro-$\ell$-quotient of $G_K$. Since $\mu_\ell\subset K$, the action of $G_k$ on the pro-$\ell$-Sylow group of the inertia group of $w$ factors through $G_k(\ell)$. We see that the canonical projection $G_K\to G_K(\ell)$ is an isomorphism. Therefore we have:
\[ \mathrm{H}^3(K,\mathbb{Z}/\ell\mathbb{Z}) \cong\mathrm{H}^3(G_K(\ell),\mathbb{Z}/\ell\mathbb{Z}) \cong\mathrm{H}^3(G_\ell,\mathbb{Z}/\ell\mathbb{Z}) \]
Now the equivalence of (i) and (ii) follows from [Se], Ch. I, §4.
